A performance assessment of evolutionary algorithms in a volunteer computing system

### 2016-03-13 PPSN2016/paper.Rnw
Explaining the different experiments

experiment.aggregates <- data.frame(Experiment=character(),
                                    milliseconds=character(),
                                    IPs=character(),
                                    actualIPs = character(),
                                    PUTs = character(),
                                    actualPUTs = character(),
                                    count = character())

@@

    experiment.aggregates <- rbind( experiment.aggregates,
                                   data.frame( Experiment=i,                   
                                              milliseconds=median(as.integer(this.data$milliseconds)),
                                              IPs=median(as.integer(this.data$IPs)),
                                              actualIPs=median(as.integer(this.data$actualIPs)),
                                              PUTs=median(as.integer(this.data$PUTs)),
                                              actualPUTs=median(as.integer(this.data$actualPUTs)),
                                              count= length(this.data$IPs))
                                   )

@@

for ( i in these.experiments ) {
    this.data <- read.csv(paste0("../data/2016-PPSN/ips-reboots-cache=32-",i,".csv"))
    this.len <- length(this.data$PUTs)
    seq <- 1:this.len
    ips.data <- rbind(ips.data,
                      data.frame(Experiment=rep(i,length(this.data$PUTs)),
                                 idu=seq,
                                 idx=seq/this.len,
                                 PUTs=as.integer(this.data$PUTs),
                                 reboots=as.integer(this.data$reboots))
                       )

    ips.aggregates <- rbind( ips.aggregates,
                            data.frame(Experiment=i,
                                       PUTs = median(this.data$PUTs),
                                       reboots = median(this.data$reboots),
                                       avgPUTs = mean(this.data$PUTs),
                                       avgreboots = mean(this.data$reboots),
                                       totalIPs =this.len )
                            )
}
@

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
Experiment & Median Time (ms) & \multicolumn{2}{|c|}{Median IPs} & \multicolumn{2}{|c|}{Median PUTs} & \#Runs\\
           &           & Total & Contributing & Total & Contributing & \\
\hline
base & 120056.0 & 10 & 4 & 256.0 & 21.0 & 81\\
\hline
graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36\\
graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36\\
\hline
reboot & 513354.5 & 9 & 6 & 545.5 & 62.5 & 40\\
\hline
cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 55\\
\hline
\end{tabular}
%% Initially generated with this code, 
%% <<experiments, cache=FALSE,echo=FALSE>>=
%% experiment.aggregates$Experiment <- c("base","graph1","graph2","graph3","reboot","cachereboot")
%% kable(experiment.aggregates)
%% @ %def

\end{center}
\end{table}
%
Since the point of using volunteer computing system is to cut costs,
all experiments were set up using the OpenShift
PaaS, which provides a free tier, making the whole experiment cost
equal to \$0.00. In fact, the NodIO system can be deployed to any
system as long as it can run node.js and access to the filesystem, for
log files, is available. All these logs are available under
request. Processed files are available in the same repository that
hosts this paper, as well as the scripts needed to process them.

@@

The baseline, called {\em base} in the

>>>>

experiment.aggregates <- data.frame(Experiment=character(),
                                    milliseconds=character(),
                                    IPs=character(),
                                    actualIPs = character(),
                                    PUTs = character(),
                                    actualPUTs = character(),
                                    count = character(),
                                    totalIPs = character())

@@

    this.ip.data <- read.csv(paste0("../data/2016-PPSN/ips-reboots-cache=32-",i,".csv"))
    this.len <- length(this.ip.data$PUTs)
    seq <- 1:this.len
    ips.data <- rbind(ips.data,
                      data.frame(Experiment=rep(i,length(this.ip.data$PUTs)),
                                 idu=seq,
                                 idx=seq/this.len,
                                 PUTs=as.integer(this.ip.data$PUTs),
                                 reboots=as.integer(this.ip.data$reboots))
                       )

    ips.aggregates <- rbind( ips.aggregates,
                            data.frame(Experiment=i,
                                       PUTs = median(this.ip.data$PUTs),
                                       reboots = median(this.ip.data$reboots),
                                       avgPUTs = mean(this.ip.data$PUTs),
                                       avgreboots = mean(this.ip.data$reboots),
                                       totalIPs =this.len )
                            )

    experiment.aggregates <- rbind( experiment.aggregates,
                                   data.frame( Experiment=i,                   
                                              milliseconds=median(as.integer(this.data$milliseconds)),
                                              IPs=median(as.integer(this.data$IPs)),
                                              actualIPs=median(as.integer(this.data$actualIPs)),
                                              PUTs=median(as.integer(this.data$PUTs)),
                                              actualPUTs=median(as.integer(this.data$actualPUTs)),
                                              count= length(this.data$IPs),
                                              totalIPs=this.len)
                                   )

@@

@

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Experiment & Median Time & \multicolumn{2}{|c|}{Median IPs} &
\multicolumn{2}{|c|}{Median PUTs} & \#Runs & Different IPs\\
           &   (ms)        & Total & Used & Total & Used &
           &\\
\hline
base & 120056.0 & 10 & 4 & 256.0 & 21.0 & 81 & 262\\
\hline
graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36 & 96\\
graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36 & 73\\
\hline
reboot & 796214.0 & 9 & 6 & 646.0 & 67.0 & 45 & 168\\
\hline
cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 55 & 225\\
\hline
\end{tabular}
%% Initially generated with this code, 
%% <<experiments, cache=FALSE,echo=FALSE>>=
%% experiment.aggregates$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
%% kable(experiment.aggregates)
%% @ %def

\end{center}
\end{table}
%
\begin{figure}[htbp]
\centering

<<zipf, cache=FALSE,echo=FALSE,fig.height=3>>=
ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))
@ %def graph for IPs and puts

\caption{Ranked and normalized number of contributions (PUTs) per IP,
  plotted in a log $y$ scale.}
\label{fig:zipf}
\end{figure}

Since the point of using volunteer computing system is to cut costs,
all experiments were set up using the OpenShift
PaaS, which provides a free tier, making the whole experiment cost
equal to \$0.00. In fact, the NodIO system can be deployed to any
system as long as it can run node.js and access to the filesystem, for
log files, is available. All these logs are available under
rrequest. Processed files are available in the same repository that
hosts this paper, as well as the scripts needed to process them.

@@

The baseline, called {\em base} in the Table, uses simply the
mechanism of dropping contributions from clients if they are repeated
to avoid overlap. Clients still get a random chromosome from the
server, so they can in fact proceed with the algorithm and even finish
it. The fact that these users were not contributing to the pool was
conveyed by inserting a graph in the client that represented the size
of the cache and labeled ``How am I contributing''. The user could
then realize if his contribution was being 0 and thus do something
about it; these are labeled {\tt graph1} and 2. However, a stuck
client could still add to the pool if evolution got unstuck, so we
added a mechanism for rebooting the client, that is, reloading the
page, if the cache size was found to be less than 1, which indicated a
run recently initiated. This batch of runs is labelled {\tt
  reboot}. However, we found that this size of the cache was actually
skipped by some slow clients or too fast evolutions, so that in the
last case, labeled {\tt cachecrash}, clients rebooted if the cache
collapsed by more than a certain length, indicating a cache that had
been voided and was filling again.
